---
title: "Attrition Classification Modeling"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

```{r}
library(dplyr)
library(ggplot2)
library(class)
library(caret)
library(e1071)
library(earth)

source('utils.R', local=knitr::knit_global())
```

## Load Datasets

```{r}
df <- read.csv('data/CaseStudy2-data.csv')
```

## Remove Irrelevant Columns from Data

Some features in the dataset don't have meaningful information for data 
analysis, such as "EmployeeNumber." These columns will be manually removed. 

```{r}
# manually remove columns that don't contain valuable data
pruned_df <- df %>%
  select(-c(EmployeeCount, EmployeeNumber, Over18, StandardHours))
```

## Data Partitioning

#### Split the Dataset into Training and Testing Sets

The train-test split procedure is used to estimate the performance of our
algorithms by calculating performance on seen and unseen data. 

```{r}
# split data into training and testing sets
train_test <- train_test_split(pruned_df, 0.70)

# assign new dataframes
train_df <- train_test$train
test_df <- train_test$test

paste0('Number of rows in training set: ', nrow(train_df))
paste0('Number of rows in testing set: ', nrow(test_df))
```

## Classification : Naive Bayes 

```{r}

```
